---
output: 
  pdf_document:
    template: template.tex
---

```{r echo = FALSE, message = FALSE}
library(tidyverse)
```

\section{Review}

```{r}
model_mpg <- lm(mpg ~ hp + am + wt, data = mtcars) 

summary(model_mpg)


anova(model_mpg) #anova stands for analysis of variance

```

\vspace{10cm}

```{r}
678.37 + 202.24 + 65.15 + 180.29

mtcars %>% 
  mutate(error = mpg - mean(mpg)) %>% 
  mutate(error_sq = error^2) %>% 
  glimpse()

mtcars %>% 
  mutate(error = mpg - mean(mpg)) %>% 
  mutate(error_sq = error^2) %>%  
  summarize(sum(error_sq))
```

\vspace{5cm}
```{r}
summary(model_mpg)$r.squared

(678.37 + 202.24 + 65.15)/(678.37 + 202.24 + 65.15 + 180.29)
```

\newpage

\section{Binomial Likelihood Example}

You have a friend who claims that they can identify caffeinated and non-caffeniated coffee. Let their success rate be an unknown parameter $p$. You do a double blind test with 10 cups of coffee. Your friend identifies 8 cups of coffee correctly. Which of the following is the most likely estimation for $p$?

a) 0.3
b) 0.5 
c) 0.8

The answer may seem intuitive but let's see the reasoning behind it


\newpage
```{r fig.align='center', fig.height=2, fig.width=3}
p <- seq(from = 0, to = 1, by = 0.001)
likelihood <- dbinom(x = 8, size = 10, prob = p)

data<- as.data.frame(cbind(p, likelihood))

data %>% 
  ggplot(aes(x = p, y = likelihood)) +
  geom_line()

data %>% 
filter(likelihood == max(likelihood))

```

```{r fig.align='center', fig.height=2, fig.width=3}
data %>% 
  ggplot(aes(x = p, y = log(likelihood))) +
  geom_line()

data %>% 
filter(log(likelihood) == max(log(likelihood)))
  
```


\newpage

Estimating $p$ is quite easy when we are only testing one friend. We are now going to consider two scenarios.

The experiment is repeated with a friend who tests 50 cups of coffee and correctly identifies 40 cups. What is the estimate for $p$?



\vspace{5cm}

In second scenario:

You repeat the experiment with five friends where each friend tests 10 cups of coffee. Two friends are correct about 7 cups, 1 friend is correct about 8 cups, 2 friends are correct about 9 cups. What is the estimate for $p$?

\newpage

\section{Binomial Likelihood Using Calculus}



