---
output: 
  pdf_document:
    template: template.tex
---

```{r echo = FALSE, message = FALSE}
library(tidyverse)
library(fivethirtyeight)
```

\fbox{\begin{minipage}{35em}
Names of Group Members: 
\vspace{2cm}

\end{minipage}}

\section{In Class Activity - Submission}

After you are done answering the questions in your groups. Write combined solutions to this sheet. Note that will __not__ receive points merely by writing down the answer. Show your work.

Q1. What is the probability that a randomly selected Bob Ross painting will have a tree in it?

\fbox{\begin{minipage}{35em}
  \hfill
\vspace{3cm}
\end{minipage}}


Q2. What is the probability that a randomly selected Bob Ross painting will have a tree and a mountain?

\fbox{\begin{minipage}{35em}
  \hfill
\vspace{3cm}
\end{minipage}}

Q3. You sat down and are watching an episode of _The Joy of Painting_. Bob Ross has already painted a mountain, what is the probability that this painting will have a tree?

\fbox{\begin{minipage}{35em}
  \hfill
\vspace{3cm}
\end{minipage}}

\pagebreak


\vspace{3cm}

Q4. Do the events of a painting having a tree and a painting having a mountain seem independent? Why or why not?

\fbox{\begin{minipage}{35em}
  \hfill
\vspace{4cm}
\end{minipage}}

\vspace{2cm}


Q5. Fill in the blanks in the code below for the plot on your handout.

```{r eval = FALSE}
bob_ross %>% 
  ggplot(aes(x = ___________________, 
             fill = _____________________________)) +
  _____________________________
```

\vspace{2cm}

Q6. Based on the data on your handout, calculate the expected household size as well as the variance.

\fbox{\begin{minipage}{35em}
  \hfill
\vspace{6cm}
\end{minipage}}
