---
output: 
  pdf_document:
    template: template.tex
---

```{r message = FALSE, echo = FALSE}
library(tidyverse)
```


\section{Continuous Random Variables}

__Definition__

\vspace{5cm}

__Examples__

\vspace{5cm}


__Probability Density Function__

\vspace{10cm}

\pagebreak
__Example__

\[
  f(x) =
  \begin{cases}
                                   12x^2(1-x) & \text{if $0\leq x \leq 1$} \\
                                   0 & \text{otherwise} 
  \end{cases}
\]

Plot the graph of $f(x)$. Is $f(x)$ a probability density function? Why or why not?


\vspace{8cm}

If $f(x)$ is a pdf then calculate the following.

a. $P(X = 0.4)$
\vspace{3cm}


b. $P(X < 0.4)$

\vspace{4cm}


c. $P(0.6 \leq X \leq 0.8)$

\vspace{5cm}


\pagebreak

Calculate E[X]

\vspace{7cm}

Calculate Var[X]

\pagebreak

\section{Review}

Q1.

```{r fig.height=2, fig.width=3}
data.frame (x = c(0, 0.5, 1),
                y = c(0, 2, 0)) %>% 
  ggplot(aes(x = x, y = y)) +
  geom_line() +
  labs(y = "f(x)")

```

Is the function depicted with the plot above a pdf? Why/not?

\vspace{4cm}

Q2. 
a. A lottery game requires players to guess six numbers from 1 to 49. If a player guessed all the six numbers then they win the lottery of the week. What is the probability that a player will win the lottery?
\vspace{3cm}

b. What is the expected number of weeks until the first win for a player?
\vspace{3cm}

c. What is the variance?

\pagebreak

Q3. A customer service agent receives on average 2 calls every 5 minutes. 

a. What is the probability that they will receive 0 calls in the next 5 minutes?
\vspace{5cm}


b. What is the probability that they will receive more than 2 calls in the next 5 minutes?
\vspace{5cm}



