---
title: "Statistical Inference for Single Mean"
author: "Dr. Dogucu"
date: "2019-11-20"
header-includes:
   - \usepackage{mathtools}
output: 
  xaringan::moon_reader:
    css: slide-styles.css
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. All Rights Reserved.</div> 

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center")
options(scipen=0)

```

```{r message = FALSE, echo = FALSE}
library(tidyverse)
library(nycflights13)
```

---
## Review

An eight-month pregnant person takes a pregnancy test and the test yields a negative result. Which of the following is true? The test result is a Type II error. True or False.

---
## Review

Among Americans who own landlines, you want to know whether majority or minority of them will pick up their phones. You call 30 random numbers and see that 40% of those numbers pick up. Test your hypotheses at $\alpha = 0.05$

---
## Review

<table class="tg">
  <tr>
    <th class="tg-lboi">z </th>
    <th class="tg-lboi">p-value</th>
    <th class="tg-lboi">Decision</th>
  </tr>
  <tr>
    <td class="tg-lboi">|z|&gt; critical value</td>
    <td class="tg-lboi">p-value &lt; $\alpha$ </td>
    <td class="tg-lboi">Reject $H_0$</td>
  </tr>
  <tr>
    <td class="tg-lboi">|z|&lt; critical value</td>
    <td class="tg-lboi">p-value &gt; $\alpha$</td>
    <td class="tg-lboi">Fail to reject $H_0$</td>
  </tr>
</table>

---
## Review

Your friend also is interested in the same question. They call 100 random numbers and see that 40% of those numbers pick up. Test their hypotheses at $\alpha = 0.05$
---
## Review

Note that pretty much all opinion surveys (Gallup etc.) we have seen before, had large sample sizes thus the standard errors were really small, the z-values were reallly large and p-values were really small.
---
## Note

One sided tests are less common since we do not know whether the sample statistic will be less than or more than the population parameter we want to compare to.

---

## Conditions

We can rely on Central Limit Theorem in order to make inference for a single mean if the following conditions have been met 

__Independence__ The sample data must be independent

__Normality__ If the sample size is small, then the sample data must come from a population with a normal distribution. 

---
## Checking Normality

Normality can be a difficult condition to test since "small" is a relative term so we use a rule of thumb

- If $n < 30$ and there are no outliers in the data then we assume the data came from a population with a normal distribution. 

- If $n \geq 30$ and there are no outliers then we assume the sampling distribution of sample means is nearly normal even if the population does not follow a normal distribution.

---
## CLT

If the conditions are met then

$\bar x \sim N(\mu, \frac{\sigma^2}{n})$

---

## CLT

If the conditions are met then

$\bar x \sim N(\mu, \frac{\sigma^2}{n})$

Houston, we have a problem! We do not know $\sigma$ so we may not calculate standard error easily. 

---

## How did we handle this problem before?

When making inference for single proportion and difference of two proportions

$\hat p \sim N(p, \frac{p(1-p)}{n})$

$(\hat p_1 - \hat p_2) \sim N((p_1 - p_2), \frac{p_1(1-p_1)}{n_1} + \frac{p_2(1-p_2)}{n_2})$

we did not know the population proportions. How did we deal with this problem when calculating standard error?
---
## How did we handle this problem before?

We used the sample proportions. 
---
## How can we handle our current problem?


Since we do not know $\sigma$, we will instead use $s$.

Also, rather than using Normal distribution, we will now rely on $t$-distribution.

For instance when we calculate a confidence interval would be

$\bar x \pm t^*_{df} \times \frac{s}{\sqrt n}$

---
## t-distribution

```{r}
ggplot(data = data.frame(x = c(-3, 3)), aes(x)) +
  stat_function(fun = dt, n = 101, args = list(df = 2)) + 
  scale_y_continuous(breaks = NULL) +
  ylab("f(x)") +
  theme(text = element_text(size = 20))

```

---

## t-distribution vs Normal Distribution

```{r}

ggplot(data = data.frame(x = c(-3, 3)), aes(x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 1), aes(color = "z"))  +
  stat_function(fun = dt, n = 101, args = list(df = 1), aes(color = "t")) + 
  ylab("f(x)") +
  scale_y_continuous(breaks = NULL) +
  scale_colour_manual(name="Distribution",
  values=c(z="red", t="blue")) +
  theme(text = element_text(size = 20))


```
---

## t-distribution

When we learned normal distribution we looked into two parameters its mean which signified the center of the distribution and its variance which signified its spread. 

In $t$-distribution, we will use one parameter only which is named degrees of freedom. 

df = n - 1
---
## t-distribution with different df

```{r}

ggplot(data = data.frame(x = c(-3, 3)), aes(x)) +
  stat_function(fun = dt, n = 101, args = list(df = 1), aes(color = "df1")) + 
  stat_function(fun = dt, n = 101, args = list(df = 2), aes(color = "df2")) + 
  stat_function(fun = dt, n = 101, args = list(df = 4), aes(color = "df4")) + 
  stat_function(fun = dt, n = 101, args = list(df = 7), aes(color = "df7")) + 
  stat_function(fun = dt, n = 101, args = list(df = 10), aes(color = "df9")) + 
  ylab("f(x)") +
  scale_y_continuous(breaks = NULL) +
  scale_colour_manual(name="df",
  values=c(df1 = "red", df2="blue", df4 = "green", df7 = "black", df9 = "orange")) +
  theme(text = element_text(size = 20))


```

---
## t-distribution vs Normal Distribution

Note that as degrees of freedom increases t-distribution approaches a normal distribution. 
---
## t-distribution using R

```{r fig.height=3}
x <- seq(-4,4, 0.01)
y <- dt(x, df = 1)
data <- data.frame(x = x, y = y)
qplot(x, y, data = data, 
      geom = "line")+
  geom_ribbon(data = subset(data, x >- 4 & x < -2),
              aes(ymax = y), ymin = 0,
              fill = "red", colour = NA, alpha = 0.5)+
  scale_y_continuous(limits = c(0, .4)) +
  ylab("f(x)")

```



```{r echo = TRUE}
pt(q = -2, df = 1) # P(t < -2)

qt(0.1475836, df = 1)
```

---

## Finding critical values

Assume that we want to calculate a 95% confidence interval for a sample mean from a sample with $n = 15$. What is the critical $t$ value?

```{r fig.height=3}
x <- seq(-4,4, 0.01)
y <- dt(x, df = 14)
data <- data.frame(x = x, y = y)
qplot(x, y, data = data, 
      geom = "line")+
  geom_ribbon(data = subset(data, x >- 4 & x < -2.144787),
              aes(ymax = y), ymin = 0,
              fill = "red", colour = NA, alpha = 0.5)+
  geom_ribbon(data = subset(data, x >2.144787 & x < 4),
              aes(ymax = y), ymin = 0,
              fill = "red", colour = NA, alpha = 0.5)+
  scale_y_continuous(limits = c(0, .4)) +
  ylab("f(x)") +
  theme(text = element_text(size = 20))

```

```{r echo = TRUE}
qt(0.025, df = 14)
```

---
## Confidence Interval for Single Mean

You might have seen $\pm$ written many times on snacks that you eat. Many companies weigh the items they produce but they do not necessarily weigh all the items. They take a sample of items. A chocolate company measured the weight of 20 chocolate bars that they produced and reported the 95% CI for the mean weight as $(3.1, 3.9)$ oz. How would this confidence interval get reported on the packaging of the chocolate bar? What is the sample mean and standard deviation of the weight of chocolate bars? 

---
## Exercise 2 

```{r echo = TRUE}
set.seed(123)

flights %>%
  filter(dest == "SNA") %>% 
  sample_n(30) %>% 
  summarize(mean_delay = mean(arr_delay, na.rm = TRUE), 
            sd_delay = sd(arr_delay, na.rm = TRUE))
```

For a sample of flights leaving NYC airports and arriving in SNA, on average planes arrive 3.33 minutes early with sd = 27.2.
---
## Exercise 2 

For flights arriving in SNA is the average arrival delay significantly different than 0? Assume the conditions have been met and set $\alpha$ to 0.05. 


---

## More exercises 

From OpenIntro: 7.12, 6.24, 7.14, 6.8