---
title: "Introduction to Probability"
author: "Dr. Dogucu"
date: "2019-10-09"
output: 
  xaringan::moon_reader:
    css: slide-styles.css
      
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"> 
 Copyright &copy; <a href="https://mdogucu.ics.uci.edu">Dr. Mine Dogucu</a>. All Rights Reserved.</div> 

---

```{r message = FALSE, echo = FALSE}
library(tidyverse)
library(titanic)
```

## Announcements

- No reading questions for Monday

---
## Summarizing Sleep Data by Hand

```{r echo = FALSE, warning = FALSE, message = FALSE}
you_sleep <- read_csv("data/you-sleep.csv")
you_sleep$sleep
```

```{r eval = FALSE, echo = FALSE}
sleep <- c(7,8,8,8,8,7,8,4,7,7.5)
you_sleep <- tibble(sleep = sleep)
you_sleep %>% 
  summarize(mean(sleep), median(sleep),
            quantile(sleep, 0.25), 
            quantile(sleep, 0.75),
            var(sleep),
            sd(sleep))

you_sleep %>% 
  ggplot(aes(x = "", y = "sleep")) +
  geom_boxplot()
```

Calculate 1)) variance, 2)standard deviation, 
---

## Mean (1 or 3) and Std Dev (1 or 3)

```{r echo = FALSE, warning = FALSE, message = FALSE, fig.align='center'}
set.seed(12345)
mean3_sd1 <- rnorm(1000, 3, 1)
mean1_sd3 <- rnorm(1000, 1, 3)
mean1_sd1 <- rnorm(1000, 1, 1)
mean3_sd3 <- rnorm(1000, 3, 3)

response <- c(mean3_sd1, 
                  mean1_sd3, 
                  mean1_sd1, 
                  mean3_sd3)

plot_type <- c(rep("A",1000),
               rep("B",1000),
               rep("C",1000),
               rep("D",1000))


mean_sd <- tibble(response = response,
                  plot_type = plot_type)

mean_sd %>% 
  ggplot(aes (x = response)) +
  geom_histogram() +
  facet_wrap(.~plot_type) +
  labs(x = "", title = "") +
  theme(text = element_text(size=30))




```

---

## Descriptives Notation

| Description        	| Population Parameter 	| Sample Statistic 	|
|--------------------	|----------------------	|------------------	|
| Mean               	|         $\mu$       	|     $\bar{x}$    	|
| Variance           	|      $\sigma^2$      	|       $s^2$      	|
| Standard Deviation 	|       $\sigma$       	|        $s$       	|


---

## Definitions

---
## Writing about Probability Using Set Notation
---
## Axioms of Probability


1) Each probability must be between 0 (inclusive) and 1 (inclusive)

$0 \leq P(A) \leq 1$ where $A \subset S$

2) The total of probabilities must be 1. 

$P(S) = 1$

3) The probability of disjoint events is additive. 

$P(A \cup B) = P(A) + P(B)$

## Venn Diagrams


## Within One Week

* Due Friday:  
 + HW  (Introduction to Data) 
 Note that your new homework will be posted on this Friday. Make sure to check the course website.
* Due Monday:   
  + No reading question 
* Due Wednesday:
  + Reading question

---
